Plots

### 2025-10-17 src/plots.Rmd
Use bar plots instead of lines and symbols, reorganized plots

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "G5" = "GPT-5 (G5)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_shapes = 21:26,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_add = c("R = Range"),
        ...
    ) {

@@

@@

# Model full name
gpt3_name <- "G3"
gpt4_name <- "G4"
gpt5_name <- "G5"
isva_name <- "ISV"
iva5_name <- "I5"

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    text_mid_vjust = 2,
    sort_group = F,
    x_limits = c(0.2, 0.9),
    x_expand = waiver(),
    text_scale = 1.25
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult, 12+ Years (n=", n_adult, ")"),
    text_mid_vjust = 2.5,
    text_scale = 1
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child, 28 Days to 11 Years (n=", n_child, ")"),
    text_mid_vjust = 4,
    text_scale = 1
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_mid_vjust = 3.5,
    text_scale = 1.25
)

>>>>

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.1),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "G5" = "GPT-5 (G5)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_shapes = 21:26,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_add = c("R = Range"),
        ...
    ) {

@@

# Plot performance of models using grouped bars
plot_perf_bar <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        x_hpos = -0.23 * text_scale,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = ((x_hpos * -1) + 0.02) * text_scale),
        x_shapes = 21:26,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        x_color_size = 3 * text_scale,
        y_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_key_size = 0.5,
        ...
    ) {

    # Get custom colors for models
    default_colors <- scales::hue_pal()(length(unique(df_plot[[color]])))
    names(default_colors) <- c(unique(as.character(df_plot[[color]])))

    # Format plot data
    df_plot <- df_plot %>%
        mutate(
            x = metric,
            y = .data[[group]],
            color = .data[[color]]
        )

    # Sort bars within groups (ascending)
    df_pos <- df_plot %>%
        group_by(y) %>%
        arrange(x, .by_group = TRUE) %>%  
        mutate(
            rank_in_group = row_number(),
            n_colors = n()
        ) %>%
        ungroup()

    # Reverse sort groups by highest-performing bar
    group_levels_order <- df_pos %>%
        group_by(y) %>%
        summarize(max_x = max(x)) %>%
        arrange(max_x) %>%
        pull(y)

    df_pos <- df_pos %>%
        mutate(
            y = factor(y, levels = group_levels_order),
            group_center = as.numeric(y)
        )

    # Bar width and offset
    group_width <- 0.8
    max_items <- max(df_pos$n_colors)
    bar_width <- group_width / max_items

    df_pos <- df_pos %>%
        mutate(
            offset = (rank_in_group - (n_colors + 1)/2) * bar_width,
            x_pos = group_center + offset,
            label_text = paste0(color, ": ", format(round(x, 2), nsmall = 2)),
            label_x = x_pos
        )

    # Identify top-performing model in each group
    df_pos <- df_pos %>%
        group_by(y) %>%
        mutate(
            top_model = x == max(x),
            fontface = ifelse(top_model, "bold", "plain"),
            alpha_val = ifelse(top_model, 1, 0.35),       # transparency for non-top bars
            outline_color = ifelse(top_model, "black", NA) # black outline for top bars
        ) %>%
        ungroup()

    # Compute range and modify group labels using square brackets
    group_labels <- df_pos %>%
        group_by(y) %>%
        summarize(range_x = max(x) - min(x)) %>%
        mutate(label = paste0(as.character(y), "\n[Range=", format(round(range_x, 2), nsmall = 2), "]")) %>%
        arrange(match(y, levels(df_pos$y)))

    # Replace factor levels with new labels including range
    levels(df_pos$y) <- group_labels$label

    # Plot metric for each model
    out <- ggplot(df_pos, aes(x = x_pos, y = x, fill = color)) +
        geom_col(
            width = bar_width * 0.9,
            position = "identity",
            color = df_pos$outline_color,
            alpha = df_pos$alpha_val
        ) +
        geom_text(
            aes(
                x = label_x,
                label = label_text,
                fontface = fontface
            ),
            y = x_hpos,
            hjust = 0,
            vjust = 0.5,
            size = x_color_size
        ) +
        coord_flip() +
        scale_fill_manual(
            values = default_colors
        ) +
        scale_x_continuous(
            breaks = seq_along(levels(df_pos$y)),
            labels = levels(df_pos$y),
            expand = c(0, 0)
        ) +
        scale_y_continuous(
            expand = x_expand,
            breaks = x_breaks,
            limits = x_limits
        ) +
        labs(
            title = title,
            x = y_title,
            y = x_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            legend.key.size = unit(legend_key_size, "lines"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)
}

@@

# Model full name
gpt3_name <- "GPT-3.5"
gpt4_name <- "GPT-4"
gpt5_name <- "GPT-5"
isva_name <- "InSilicoVA"
iva5_name <- "InterVA-5"

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf_bar(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25),
    text_scale = 0.95
)

@@

# Plot adult perf for pccc by cod
fig_pccc_cod_adult <- plot_perf_bar(
    df_long %>% filter(age == "adult"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Adult, 12+ Years (n=", n_adult, ")"),
    text_scale = 0.7,
    x_hpos = -0.15,
    x_expand = expansion(add = 0.15)
)

@@

# Plot child perf for pccc by cod
fig_pccc_cod_child <- plot_perf_bar(
    df_long %>% filter(age == "child"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Child, 28 Days to 11 Years (n=", n_child, ")"),
    text_scale = 0.92,
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25)
)

@@

# Plot neo perf for pccc by cod
fig_pccc_cod_neo <- plot_perf_bar(
    df_long %>% filter(age == "neo"),
    calc_pccc,
    group = "cod_name",
    title = paste0("Neonatal, <28 Days (n=", n_neo, ")"),
    x_title = "Partial Chance Corrected Concordance (PCCC)",
    text_scale = 0.92,
    x_hpos = -0.24,
    x_expand = expansion(add = 0.25)
)
